# Datasets {#data}

## International Stroke Trial Example Dataset {#ist}

The International Stroke Trial database contains individual patient data
from the International Stroke Trial (IST), a multi-national randomized trial 
conducted between 1991 and 1996 (pilot phase between 1991 and 1993) that aimed 
to assess whether early administration of aspirin, heparin, both or neither 
influenced the clinical course of acute ischaemic stroke 
[@sandercock1997international]. The IST dataset includes data on 19,435 patients 
with acute stroke, with 99\% complete follow-up. De-identified data are 
available for download at https://datashare.is.ed. ac.uk/handle/10283/128. This 
study is described in more detail at the bottom of this page, and in the 
corresponding block quote reference. In the example data for this workshop, we 
consider a sample of 5,000 patients and the binary outcome of recurrent ischemic 
stroke within 14 days after randomization. Also in our example data, we ensure 
that we have subjects with a missing outcome. 

```{r load_ist_data_intro, message=FALSE, warning=FALSE}
library(tidyverse)

# read in data
ist <- read_csv("https://raw.githubusercontent.com/tlverse/deming2019-workshop/master/data/ist_sample.csv")
ist
```

For the purposes of this workshop, we we start by treating the data as
independent and identically distributed (i.i.d.) random draws from a very large
target population. We could, with available options, account for the clustering
of the data (within sampled geographic regions), but, for simplification, we 
avoid these details in these workshop presentations, although modifications of 
our methodology for biased samples, repeated measures, etc., are available.

We have 26 variables measured, of which 1 variable is set to be the outcome of
interest. This outcome, $Y$, indicates recurrent ischemic stroke within 14 days 
after randomization (`DRSISC` in `ist`); the treatment of interest, 
$A$, is the randomized aspirin vs. no aspirin treatment allocation (`RXASP` in 
`ist`); and the adjustment set, $W$, consists simply of other variable measured 
at baseline. In this data our outcome is occasionally missing, but we do not 
need to create a variable indicating this missingness (such as $\Delta$) for 
analyses in the `tlverse`. If we let $\Delta$ denote the indicator that 
the outcome is missing such that $\Delta = 1$ when the outcome is observed and 
$\Delta = 0$ when the outcome is not observed, then we can denote our observed 
data structure as $n$ i.i.d. copies of $O_i = (W_i, A_i, \Delta_i, \DeltaY_i)$, 
for $i = 1, \ldots, n$.

Using the [`skimr` package](https://CRAN.R-project.org/package=skimr), we can
quickly summarize the variables in our data:

```{r skim_ist_data, message=FALSE, warning=FALSE}
library(skimr)
skim(ist)
```

A convenient summary of the relevant variables is given just above.

> "The study had a prospective, randomised, open treatment, blinded outcome 
> (PROBE) design. The inclusion criteria were: clinical diagnosis of acute 
> ischaemic stroke, with onset within the previous 48 hours and no clear 
> indication for, or clear contraindication to, treatment with aspirin or 
> subcutaneous heparin. Unlike many stroke trials of that era (and 
> subsequently), the study did not set an upper age limit. Patients were to 
> have a CT brain scan to confirm the diagnosis of stroke, and this was to be 
> done before randomisation if at all possible. To enter a patient in the 
> study, the clinician telephoned a central randomisation service (at the 
> Clinical Trial Service Unit, Oxford) during this telephone call, the baseline 
> variables were entered and checked, and once validated, the computer 
> allocated the treatment and the telephonist then informed the clinician. The 
> patients and treating clinicians were not blinded to the treatment given. 
> Early outcome data were collected by the treating physician who completed a 
> follow-up form at 14 days, death or hospital discharge (whichever occurred 
> first). This form recorded data on events in hospital within 14 days, and the 
> doctor's opinion on the final diagnosis of the initial event that led to 
> randomisation. These unblinded data, may therefore be subject to some degree 
> of bias. The primary outcome was the proportion of patients who were either 
> dead or dependent on other people for activities of daily living at six 
> months after randomisation. This outcome was collected by postal 
> questionnaire mailed directly to the patient, or (in Italy) by telephone 
> interview of the patient by a trained researcher, blinded to treatment 
> allocation. The primary outcome was therefore assessed - as far as 
> practicable - blind to treatment allocation and hence should be free from 
> bias. We re-checked the data set for inaccuracies and inconsistencies and 
> extracted data on the variables assessed at randomisation, and at the two 
> outcome assessment points: at 14-days after randomisation, death or prior 
> hospital discharge (whichever occurred first) and at 6-months.
>
> --- @sandercock2011international

## WASH Benefits Example Dataset {#wash}

The data come from a study of the effect of water quality, sanitation, hand
washing, and nutritional interventions on child development in rural Bangladesh
(WASH Benefits Bangladesh): a cluster-randomised controlled trial
[@luby2018effects]. The study enrolled pregnant women in their first or second
trimester from the rural villages of Gazipur, Kishoreganj, Mymensingh, and
Tangail districts of central Bangladesh, with an average of eight women per
cluster. Groups of eight geographically adjacent clusters were block-randomised,
using a random number generator, into six intervention groups (all of which
received weekly visits from a community health promoter for the first 6 months
and every 2 weeks for the next 18 months) and a double-sized control group (no
intervention or health promoter visit). The six intervention groups were:

1. chlorinated drinking water;
2. improved sanitation;
3. hand-washing with soap;
4. combined water, sanitation, and hand washing;
5. improved nutrition through counseling and provision of lipid-based nutrient
   supplements; and
6. combined water, sanitation, handwashing, and nutrition.

In the workshop, we concentrate on child growth (size for age) as the outcome of
interest. For reference, this trial was registered with ClinicalTrials.gov as
NCT01590095.

```{r load_washb_data_intro, message=FALSE, warning=FALSE}
library(tidyverse)

# read in data
dat <- read_csv("https://raw.githubusercontent.com/tlverse/tlverse-data/master/wash-benefits/washb_data.csv")
dat
```

We have 28 variables measured, of which 1 variable is set to be the outcome of
interest. This outcome, $Y$, is the weight-for-height Z-score (`whz` in `dat`);
the treatment of interest, $A$, is the randomized treatment group (`tr` in
`dat`); and the adjustment set, $W$, consists simply of *everything else*. This
results in our observed data structure being $n$ i.i.d. copies of $O_i = (W_i,
A_i, Y_i)$, for $i = 1, \ldots, n$.

Like before, we can summarize the variables measured in the WASH Benefits data 
set with `skimr`:

```{r skim_washb_data, message=FALSE, warning=FALSE}
skim(dat)
```

Note that the *asset* variables reflect socio-economic status of the study 
participants. 

## Veterans’ Administration Lung Cancer Trial Dataset {#vet}

This data corresponds to a study conducted by the US Veterans Administration. 
Male patients with advanced inoperable lung cancer were given either the 
standard therapy or a test chemotherapy. The primary goal of the study was to 
assess if the test chemotherapy improved survival. This data set has been 
published in @kalbfleisch2011statistical and it is available in the `MASS` and 
`survival` `R` packages. Time to death was recorded for 128 patients, and 9 
patients left the study before death. Various covariates were also documented 
for each patient. 

```{r load_ist_data_intro, message=FALSE, warning=FALSE}
library(tidyverse)

# read in data
vet <- read_csv("https://raw.githubusercontent.com/tlverse/deming2019-workshop/master/data/veteran.csv")
vet
```

A snapshot of the data set in shown below:

```{r skim_vet_data, message=FALSE, warning=FALSE}
skim(vet)
```