# Super (Ensemble Machine) Learning

Based on the [`sl3` `R` package](https://github.com/tlverse/sl3) by _Jeremy
Coyle, Nima Hejazi, Ivana Malenica, and Oleg Sofrygin_.

Updated: `r Sys.Date()`

## Introduction

Once the statistical estimation problem is defined, as described in the 
[The Targeted Learning Roadmap](#intro), we are ready to construct the TMLE:
an asymptotically efficient substitution estimator of this target quantity.

The first step in the estimation procedure is an initial estimate of the
data-generating distribution, or the relevant part of this distribution that is
needed to evaluate the target parameter. For this initial estimation, we use the
super learner [@vdl2007super], an important step for creating a robust
estimator.

### Super Learning

- A common task in statistical data analysis is estimator selection 
  (e.g., for prediction).
- There is no universally optimal machine learning algorithm for density
  estimation or prediction.
- For some data, one needs learners that can model a complex function.
- For others, possibly as a result of noise or insufficient sample size, a
  simple, parametric model might fit best.
- Super Learner, an ensemble learner, solves this issue, by allowing a
  combination of learners from the simplest (intercept-only) to most complex
  (neural nets, random forests, SVM, etc).
- It works by using cross-validation in a manner which guarantees that the
  resulting fit will be as good as possible, given the learners provided.
- Note: even a combination of poor learners can sometimes result in good fit.
  It is very important to have good candidates in our library, possibly
  incorporating known knowledge about the system in question.
  
#### General Overview of the Algorithm

**What is cross-validation and how does it work?**

- There are many different cross-validation schemes, designed to accommodate different 
  study designs and data structures.
- The figure below shows an example of 10-fold cross-validation.  
  
```{r cv_fig, echo = FALSE}
knitr::include_graphics("img/misc/vs.pdf")
```

**General step-by-step overview of the Super Learner algorithm:**

- Break up the sample evenly into V-folds (say V=10).
- For each of these 10 folds, remove that portion of the sample (kept out as
  validation sample) and the remaining will be used to fit learners (training
  sample).
- Fit each learner on the training sample (note, some learners will have their
  own internal cross-validation procedure or other methods to select tuning
  parameters).
- For each observation in the corresponding validation sample, predict the outcome
  using each of the learners, so if there are $p$ learners, then there would be
  $p$ predictions.
- Take out another validation sample and repeat until each of the V-sets of data
  are removed.
- Compare the cross-validated fit of the learners across all observations based
  on specified loss function (e.g., squared error, negative log-likelihood, ...)
  by calculating the corresponding average loss (risk).
- Either:

  + choose the learner with smallest risk and apply that learner to entire data
    set (resulting SL fit),
  + do a weighted average of the learners to minimize the cross-validated risk
    (construct an ensemble of learners), by

    + re-fitting the learners on the original data set, and
    + use the weights above to get the SL fit.

Note, this entire procedure can be itself cross-validated to get a consistent
estimate of the future performance of the SL fit.

```{r cv_fig2, echo = FALSE}
knitr::include_graphics("img/misc/SLKaiserNew.pdf")
``` 

**How to pick a Super Learner library?**

- A library is simply a collection of algorithms.
- The algorithms in the library should come from contextual knowledge and a large
  set of "default" algorithms.
- The algorithms may range from a simple linear regression model to multi-step 
  algorithms involving screening covariates, penalizations, optimizing tuning
  parameters, etc. 

#### Example: Super Learner In Prediction

- We observe a learning data set $X_i=(Y_i,W_i)$, for $i=1, ..., n$.
- Here, $Y_i$ is the outcome of interest, and $W_i$ is a p-dimensional 
  set of covariates.
- Our objective is to estimate the function $\psi_0(W) = E(Y|W)$.
- This function can be expressed as the minimizer of the expected loss:
  $\psi_0(W) = \text{argmin}_{\psi} E[L(X,\psi(W))]$.
- Here, the loss function is represented as $L$ (e.g., squared error loss,
  $L: (Y-\psi(W))^2)$).
  
#### Why use the Super Learner?

- For prediction, one can use the cross-validated risk to empirically determine
  the relative performance of SL and competing methods. 
- When we have tested different algorithms on actual
  data and looked at the performance (e.g., MSE of prediction), never does one
  algorithm always win (see below).
- Below shows the results of such a study, comparing the fits of several different learners, 
  including the SL algorithms.

```{r cv_fig3, echo = FALSE}
knitr::include_graphics("img/misc/ericSL.pdf")
```

- Super Learner performs asymptotically as well as best possible weighted 
  combination. 
- By including all competitors in the library of candidate estimators (glm, neural nets, 
  SVMs, random forest, etc.), the Super Learner will asymptotically outperform 
  any of its competitors- even if the set of competitors is allowed to grow polynomial
  in sample size. 
- Motivates the name "Super Learner": it provides a system of combining many estimators
  into an improved estimator. 

### Review of the Super Learner {-}

* Loss-function-based tool that uses V-fold cross-validation to obtain the best
  prediction of the relevant part of the likelihood that's needed to evaluate
  target parameter.

* Requires expressing the estimand as the minimizer of an expected loss, and
  proposing a library of algorithms ("learners" in `sl3` nomenclature) that we
  think might be consistent with the true data-generating distribution.

* The *discrete super learner*, or cross-validated selector, is the algorithm in
  the library that minimizes the V-fold cross-validated empirical risk.

* The *super learner* is a weighted average of the library of
  algorithms, where the weights are chosen to minimize the V-fold
  cross-validated empirical risk of the library. Restricting the weights
  ("metalearner" in `sl3` nomenclature) to be positive and sum to one (convex
  combination) has been shown to improve upon the discrete super learner
  [@polley2010super; @vdl2007super].
  
* Proven to be asymptotically as accurate as the best possible prediction
  algorithm that is tested [@vdl2003unified; @van2006oracle].  

* This background material is described in greater detail in the accompanying
  `tlverse` handbook [`sl3`
  chapter](https://tlverse.org/tlverse-handbook/ensemble-machine-learning.html). 
  
## `sl3` "Microwave Dinner" Implementation

We begin by illustrating the core functionality of the super learner algorithm
as implemented in `sl3`. For those who are interested in the internals 
of `sl3`, see this [`sl3` introductory 
tutorial](https://tlverse.org/sl3/articles/intro_sl3.html). 

The `sl3` implementation consists of the following steps:

0. Load the necessary libraries and data
1. Define the machine learning task
2. Make a super learner by creating library of base learners and a metalearner
3. Train the super learner on the machine learning task
4. Obtain predicted values

### WASH Benefits Study Example {-}

Using the IST data, we are interested in predicting recurrent stroke `DRSISC` 
using the available covariate data.

### 0. Load the necessary libraries and data {-}

```{r setup, message=FALSE, warning=FALSE}
library(kableExtra)
library(knitr)
library(skimr)
library(tidyverse)
library(data.table)
library(sl3)
library(SuperLearner)
library(origami)
set.seed(7194)
# load data set and take a peek
ist_data <- read.csv("https://raw.githubusercontent.com/tlverse/deming2019-workshop/master/data/ist_sample.csv")

head(ist_data) %>%
  kable(digits = 4) %>%
  kable_styling(fixed_thead = T, font_size = 10) %>%
  scroll_box(width = "100%", height = "250px")
```

### 1. Define the machine learning task {-}

To define the machine learning **"task"** (predict stroke `DRSISC` using the 
available covariate data), we need to create an `sl3_Task` object.

The `sl3_Task` keeps track of the roles the variables play in the
machine learning problem, the data, and any metadata (e.g., observational-level
weights, id, offset).

We are not interested in predicting missing outcomes. We set 
`drop_missing_outcome = TRUE` when we create the task. In the next chapter, we 
estimate this missingness mechanism and account for it in the estimation. 
```{r task}
# specify the outcome and covariates
outcome <- "DRSISC"
covars <- colnames(ist_data)[-which(names(ist_data) == outcome)]
# create the sl3 task
ist_task <- make_sl3_Task(
  data = ist_data,
  covariates = covars,
  outcome = outcome, 
  drop_missing_outcome = TRUE
)
# examine the task
ist_task
```

### 2. Make a super learner {-}

Now that we have defined our machine learning problem with the task, we are
ready to **"make"** the super learner. This requires specification of

* Base learning algorithms, to establish a library of learners that we think
might be consistent with the true data-generating distribution.
* Metalearner, to ensemble the base learners.

We might also incorporate

* Feature selection, to pass only a subset of the predictors to the algorithm.
* Hyperparameter specification, to tune base learners.

Learners have properties that indicate what features they support. We may use
`sl3_list_properties()` to get a list of all properties supported by at least
one learner.

```{r list-properties, message=FALSE, warning=FALSE}
sl3_list_properties()
```
Since we have a binomial outcome, we may identify the learners that support
this outcome type with `sl3_list_learners()`.

```{r list-learners, message=FALSE, warning=FALSE}
sl3_list_learners(c("binomial"))
```

Now that we have an idea of some learners, we can construct them using the
`make_learner` function.

```{r baselearners, message=FALSE, warning=FALSE}
# choose base learners
lrnr_mean <- make_learner(Lrnr_mean)
lrnr_glm <- make_learner(Lrnr_glm)
lrnr_gam <- make_learner(Lrnr_gam)
```
We can customize learner hyperparameters to incorporate a diversity of different
settings. Documentation for the learners and their hyperparameters can be found
in the [`sl3` Learners
Reference](https://tlverse.org/sl3/reference/index.html#section-sl-learners).

```{r extra-lrnr-awesome, message=FALSE, warning=FALSE}
lrnr_ranger50 <- make_learner(Lrnr_ranger, num.trees = 50)
lrnr_hal_simple <- make_learner(Lrnr_hal9001, max_degree = 2, n_folds = 2)
lrnr_lasso <- make_learner(Lrnr_glmnet) # alpha default is 1
lrnr_ridge <- make_learner(Lrnr_glmnet, alpha = 0)
lrnr_elasticnet <- make_learner(Lrnr_glmnet, alpha = .5)
```

We can also include learners from the `SuperLearner` `R` package.

```{r extra-lrnr-woah, message=FALSE, warning=FALSE}
lrnr_bayesglm <- Lrnr_pkg_SuperLearner$new("SL.bayesglm")
```

Here is a fun trick to create customized learners over a grid of parameters.

```{r extra-lrnr-mindblown, eval = FALSE}
# I like to crock pot my super learners
grid_params <- list(cost = c(0.01, 0.1, 1, 10, 100, 1000),
                    gamma = c(0.001, 0.01, 0.1, 1),
                    kernel = c("polynomial", "radial", "sigmoid"),
                    degree = c(1, 2, 3))
grid <- expand.grid(grid_params, KEEP.OUT.ATTRS = FALSE)
params_default <- list(nthread = getOption("sl.cores.learners", 1))
svm_learners <- apply(grid, MARGIN = 1, function(params_tune) {
  do.call(Lrnr_svm$new, c(params_default, as.list(params_tune)))})

grid_params <- list(max_depth = c(2, 4, 6, 8),
                    eta = c(0.001, 0.01, 0.1, 0.2, 0.3),
                    nrounds = c(20, 50))
grid <- expand.grid(grid_params, KEEP.OUT.ATTRS = FALSE)
params_default <- list(nthread = getOption("sl.cores.learners", 1))
xgb_learners <- apply(grid, MARGIN = 1, function(params_tune) {
  do.call(Lrnr_xgboost$new, c(params_default, as.list(params_tune)))})
```

Did you see `Lrnr_caret` when we called `sl3_list_learners(c("binomial"))`? 
All we need to specify is the algorithm to use, which is passed as `method` to 
`caret::train()`. The default method for parameter selection criterion with 
is set to "CV" instead of the `caret::train()` default `boot`. The summary 
metric to used to select the optimal model is `RMSE` for continuous outcomes 
and `Accuracy` for categorical and binomial outcomes.

```{r carotene, eval = FALSE}
# I have no idea how to tune a neural net (or BART machine..)
lrnr_caret_nnet <- make_learner(Lrnr_caret, algorithm = "nnet")
lrnr_caret_bartMachine <- make_learner(Lrnr_caret, algorithm = "bartMachine", 
                                       method = "boot", metric = "Accuracy",
                                       tuneLength = 10)
```

In order to assemble the library of learners, we need to **"stack"** them
together.

A `Stack` is a special learner and it has the same interface as all
other learners. What makes a stack special is that it combines multiple learners
by training them simultaneously, so that their predictions can be either
combined or compared.

```{r stack, message=FALSE, warning=FALSE}
stack <- make_learner(Stack, 
  lrnr_glm, lrnr_mean, lrnr_lasso, lrnr_gam, 
  lrnr_bayesglm, lrnr_ridge, lrnr_elasticnet
)
```

We can optionally select a subset of available covariates and pass only
those variables to the modeling algorithm.

Let's consider screening covariates based on their `randomForest` variable 
importance ranking (ordered by mean decrease in accuracy)

```{r screener}
screen_rf <- make_learner(Lrnr_screener_randomForest, nVar = 5, ntree = 100)
# which covariates are selected on the full data?
screen_rf$train(ist_task)
```

To **"pipe"** only the selected covariates to the modeling algorithm, we need to
make a `Pipeline`, which is a just set of learners to be fit sequentially, where
the fit from one learner is used to define the task for the next learner.

```{r screener-pipe}
screen_rf_pipeline <- make_learner(Pipeline, screen_rf, stack)
```
Now our learners will be preceded by a screening step.

We also consider the original `stack`, just to compare how the feature selection
methods perform in comparison to the methods without feature selection.

Analogous to what we have seen before, we have to stack the pipeline and
original `stack` together, so we may use them as base learners in our super
learner.

```{r screeners-stack, message=FALSE, warning=FALSE}
fancy_stack <- make_learner(Stack, screen_rf_pipeline, stack)
# we can visualize the stack
dt_stack <- delayed_learner_train(fancy_stack, ist_task)
plot(dt_stack, color = FALSE, height = "400px", width = "100%")
```

We will use the [default 
metalearner](https://github.com/tlverse/sl3/blob/master/R/default_metalearner.R), 
which uses [`Lrnr_solnp()`](https://github.com/tlverse/sl3/blob/master/R/Lrnr_solnp.R) 
to provide fitting procedures for a pairing of [loss 
function](https://github.com/tlverse/sl3/blob/master/R/loss_functions.R) and 
[metalearner 
function](https://github.com/tlverse/sl3/blob/master/R/metalearners.R). The
default metalearner chooses loss and metalearner pairing based on the outcome 
type. Note that any learner can be used as a metalearner.

We have made a library/stack of base learners, so we are ready to make the super 
learner. The super learner algorithm fits a metalearner on the validation-set 
predictions.

```{r make-sl, message=FALSE, warning=FALSE}
sl <- make_learner(Lrnr_sl,
  learners = fancy_stack
)
# we can visualize the super learner
dt_sl <- delayed_learner_train(sl, ist_task)
plot(dt_sl, color = FALSE, height = "400px", width = "100%")
```

We can also use `Lrnr_cv` to build a super learner, cross-validate a stack of 
learners to compare performance of the learners in the stack, or cross-validate
any single learner (see "Cross-validation" section of this [`sl3` 
introductory tutorial](https://tlverse.org/sl3/articles/intro_sl3.html)).

Furthermore, we can [Define New `sl3` 
Learners](https://tlverse.org/sl3/articles/custom_lrnrs.html) which can be used 
in all the places you could otherwise use any other `sl3` learners, including
`Pipelines`, `Stacks`, and the Super Learner. 

### 3. Train the super learner on the machine learning task {-}

Now we are ready to **"train"** our super learner on our `sl3_task` object,
`ist_task`.

```{r sl-basic, message=FALSE, warning=FALSE}
sl_fit <- sl$train(ist_task)
```

### 4. Obtain predicted values {-}

Now that we have fit the super learner, we are ready to obtain our predicted
values, and we can also obtain a summary of the results.

```{r sl-basic-summary, message=FALSE, warning=FALSE}
sl_preds <- sl_fit$predict()
head(sl_preds)
sl_fit$print()
```
## Extensions

### Cross-validated Super Learner

We can cross-validate the super learner to see how well the super learner
performs on unseen data, and obtain an estimate of the cross-validated risk of
the super learner.

This estimation procedure requires an "external" layer of cross-validation,
also called nested cross-validation, which involves setting aside a separate
holdout sample that we don’t use to fit the super learner. This
external cross-validation procedure may also incorporate 10 folds, which is the
default in `sl3`. However, we will incorporate 2 outer/external folds of
cross-validation for computational efficiency.

We also need to specify a loss function to evaluate super learner.
Documentation for the available loss functions can be found in the [`sl3` Loss
Function Reference](https://tlverse.org/sl3/reference/loss_functions.html).

``````{r CVsl, message=FALSE, warning=FALSE}
ist_task_new <- make_sl3_Task(
  data = ist_data,
  covariates = covars,
  outcome = outcome,
  drop_missing_outcome = TRUE,
  folds = make_folds(n = sum(!is.na(ist_data$DRSISC)), 
                     fold_fun = folds_vfold, 
                     V = 2)
)
CVsl <- CV_lrnr_sl(sl_fit, ist_task_new, loss_loglik_binomial)
CVsl %>%
   kable(digits = 4) %>%
   kable_styling(fixed_thead = T, font_size = 10) %>%
   scroll_box(width = "100%", height = "250px")
```

### Variable Importance Measures with `sl3`

The `sl3` `varimp` function returns a table with variables listed in decreasing
order of importance, in which the measure of importance is based on a risk
difference between the learner fit with a permuted covariate and the learner
fit with the true covariate, across all covariates.

In this manner, the larger the risk difference, the more important the variable
is in the prediction.

```{r varimp, message=FALSE, warning=FALSE}
ist_varimp <- varimp(sl_fit, loss_loglik_binomial)
ist_varimp %>%
  kable(digits = 4) %>%
  kable_styling(fixed_thead = T, font_size = 10) %>%
  scroll_box(width = "100%", height = "250px")
```

## Exercise

### Predicting Myocardial Infarction with `sl3` {#sl3ex}

Follow the steps below to predict myocardial infarction (`mi`) using the
available covariate data. We thank Prof. David Benkeser at Emory University for
making the this Cardiovascular Health Study (CHS) data accessible.

```{r ex-setup, message=FALSE, warning=FALSE}
# load the data set
db_data <-
  url("https://raw.githubusercontent.com/benkeser/sllecture/master/chspred.csv")
chspred <- read_csv(file = db_data, col_names = TRUE)
# take a quick peek
head(chspred) %>%
  kable(digits = 4) %>%
  kable_styling(fixed_thead = T, font_size = 10) %>%
  scroll_box(width = "100%", height = "200px")
```

1. Create an `sl3` task, setting myocardial infarction `mi` as the outcome and
   using all available covariate data.
2. Make a library of seven relatively fast base learning algorithms (i.e., do
   not consider BART or HAL). Customize hyperparameters for one of your
   learners. Feel free to use learners from `sl3` or `SuperLearner`. You may
   use the same base learning library that is presented above.
3. Incorporate feature selection with the `SuperLearner` screener `screen.corP`.
4. Fit the metalearning step with non-negative least squares, `Lrnr_nnls`.
5. With the metalearner and base learners, make the super learner and train it
   on the task.
6. Print your super learner fit by calling `print()` with `$`. Which learner 
   is the discrete super learner? 
7. Cross-validate your super learner fit to see how well it performs on unseen
   data. Specify `loss_squared_error` as the loss function to evaluate the
   super learner. Like above, create a new task with 2 folds of external 
   cross-validation for computational efficiency. Report the cross-validated
   mean risk of the discrete super learner and the super learner.

## Summary

* The general ensemble learning approach of super learner can be applied to a
  diversity of estimation and prediction problems that can be defined by a loss
  function.

* Plug-in estimators of the estimand are desirable because a plug-in estimator
  respects both the local and global constraints of the statistical model.

* Asymptotically linear estimators are also advantageous, since they converge to
  the estimand at $\frac{1}{\sqrt{n}}$ rate, and thereby permit formal
  statistical inference.

* If we plug in the estimator returned by super learner into the target
  parameter mapping, then we would end up with an estimator that has the same
  bias as what we plugged in. This estimator would not be asymptotically linear.

* Targeted maximum likelihood estimation (TMLE) is a general strategy that
  succeeds in constructing asymptotically linear plug-in estimators.

* In the chapters that follow, we focus on the targeted maximum likelihood
  estimator and the targeted minimum loss-based estimator, both referred to as
  TMLE.

### Exercise Solutions

Here is a potential solution to the (`sl3` Exercise -- Predicting Myocardial
Infarction with `sl3`)[@sl3ex].

```{r ex-key, eval=FALSE, message=FALSE, warning=FALSE}
chspred_task <- make_sl3_Task(
  data = chspred,
  covariates = head(colnames(chspred), -1),
  outcome = "mi"
)
glm_learner <- Lrnr_glm$new()
lasso_learner <- Lrnr_glmnet$new(alpha = 1)
ridge_learner <- Lrnr_glmnet$new(alpha = 0)
enet_learner <- Lrnr_glmnet$new(alpha = 0.5)
curated_glm_learner <- Lrnr_glm_fast$new(formula = "mi ~ smoke + beta + waist")
mean_learner <- Lrnr_mean$new() # That is one mean learner!
glm_fast_learner <- Lrnr_glm_fast$new()
ranger_learner <- Lrnr_ranger$new()
svm_learner <- Lrnr_svm$new()
xgb_learner <- Lrnr_xgboost$new()
screen_cor <- Lrnr_pkg_SuperLearner_screener$new("screen.corP")
glm_pipeline <- make_learner(Pipeline, screen_cor, glm_learner)
stack <- make_learner(
  Stack,
  glm_pipeline, glm_learner,
  lasso_learner, ridge_learner, enet_learner,
  curated_glm_learner, mean_learner, glm_fast_learner,
  ranger_learner, svm_learner, xgb_learner
)
metalearner <- make_learner(Lrnr_nnls)
sl <- Lrnr_sl$new(
  learners = stack,
  metalearner = metalearner
)
sl_fit <- sl$train(task)
sl_fit$print()
CVsl <- CV_lrnr_sl(sl_fit, chspred_task, loss_squared_error)
CVsl
```