# One-Step TMLE for Time-to-Event Outcomes

Based on the [`MOSS` `R` package](https://github.com/wilsoncai1992/MOSS)
by _Wilson Cai and Mark van der Laan_.

Updated: `r Sys.Date()`

## Learning Objectives

1. Format right-censored survival data for `MOSS`. 
2. Fit a SuperLearner initial estimate of the conditional survival function of 
failure, conditional survival function of censoring and the propensity scores 
(`initial_sl_fit`).
3. Calculate the TMLE adjustment of the conditional survival fit 
(`MOSS_hazard`).
4. Formulate a simultaneous confidence band for the estimated conditional 
survival across a range of time-points (`compute_simultaneous_ci`).

## Introduction

In this section, we explore the `MOSS` `R` package. This software performs 
ensemble machine learning with the [`SuperLearner` `R`
package](https://cran.r-project.org/web/packages/SuperLearner/index.html) and 
One-Step Targeted Maximum Likelihood Estimation (TMLE) to estimate 
counterfactual marginal survival functions and the Average Treatment Effect 
(ATE) on survival probabilities, while non-parametrically adjusting for measured 
confounding. This one-step TMLE can be executed via recursion in small local 
updates, and creates a doubly robust and semi-parametrically efficient 
estimator. Simultaneous confidence bands of the entire curve are also 
available for inference. 

### Existing Methods for Observational Survival Analysis

To facilitate comparison with other estimation procedures, the following 
additional estimators are also included in the `MOSS` `R` package:

* Inverse Censoring Probability Weighted (IPCW) estimator, which re-weights the 
  observed data by the inverse of the product of the propensity score and 
  censoring probability before applying a standard estimation method 
  [@rotnitzky2014inverse]. 
* Estimating Equation (EE) estimator, which improves IPCW by adding the sample 
  mean of the efficient influence curve and is a locally efficient and double 
  robust [@hubbard2000nonparametric]. 
 
**TMLE works well to improve the statistical efficiency of EE**
Like EE, TMLE is also doubly robust and locally efficient. In contrast to these 
two methods, TMLE performs an adjustment on the estimate of the relevant part 
of the data-generating distribution before applying the parameter mapping and 
thus always respects the parameter space (probabilities falling inside [0,1]), 
a so-called substitution/plug-in estimator. As a result, is more robust to 
outliers and sparsity than non-substitution estimators.

**Motivating one-step TMLE: monotonic survival curve**
EE and TMLE utilize efficiency theory derived for univariate parameters, making 
estimation of the survival curve a collection of univariate survival probability 
estimators. This procedure can lead to a survival curve that is not 
monotonically decreasing. The one-step TMLE implemented in `MOSS` targets the 
survival curve *as a whole* and thus ensures monotonicity, while preserving the 
desirable performance of the point-wise TMLE [@cai2019one].

## Using `MOSS` for One-Step TMLE for Time-to-Event Outcomes

`MOSS` implementation consists of the following steps:

0. Load the necessary libraries and data.
1. Specify the right-censored survival data arguments.
2. Estimate the (1) conditional survival function of failure event, (2) 
  conditional survival function of censoring event, and (3) propensity score
  with the `SuperLearner` `R` package. 
3. Perform TMLE adjustment of the initial conditional survival fit.
4. Compute standard error estimates for simulaneous inference.

### 0. Load necessary libraries and data {-}

```{r moss-0, message=FALSE, warning=FALSE}
knitr::opts_chunk(cache = TRUE)
library(MOSS)

vet_data <- read.csv("https://raw.githubusercontent.com/tlverse/deming2019-workshop/master/data/veteran.csv")
head(vet_data)
```

The variables in `vet_data` are
* `trt`: treatment type (1 = standard, 2 = test),
* `celltype`: histological type of the tumor,
* `time`: time to death or censoring in days,
* `status`: censoring status,
* `karno`: Karnofsky performance score that describes overall patient status at 
  the beginning of the study (100 = good),
* `diagtime`: months from diagnosis to randomization,
* `age`: age of patient in years, and
* `prior`: prior therapy (0 = no, 10 = yes).

### 1. Specify right-censored survival data arguments {-}

The following variables need to be specified from the data so they can 
subsequently be used as required arguments for `MOSS` functions: 

* `W`: dataframe of baseline covariates $W$.
* `A`: binary treatment vector $A$. 
* `T_tilde`: time-to-event vector $\tilde{T} = \min(T, C)$.
* `Delta`: censoring indicator vector $\Delta = I(T \leq C)$.
* `t_max`: the maximum time to estimate the survival probabilities.

We can specify these variables with our observed `vet_data`.
```{r moss-1, message=FALSE, warning=FALSE}
T_tilde <- vet_data$time
Delta <- vet_data$status
A <- vet_data$trt
W <- vet_data[, c(3, 6:9)]
t_max <- max(T_tilde)
```

### 2. Obtain initial estimates with `SuperLearner` `R` package {-}

We will use the `initial_sl_fit()` function to specify the data (as we
defined it above) and the `SuperLearner` library for initial estimation of each 
of the following components of the likelihood:

1. conditional survival function of failure event given treatment and 
  confounders,
2. conditional survival function for censoring event given treatment and 
  confounders, and
3. propensity score of treatment given confounders.

We are forgetting one component of the likelihood that requires estimation: the 
joint distribution of confounders! In `MOSS` this estimation is done for us 
under the hood, and we do not use the `SuperLearner`. Do you recall why we do 
not use the `SuperLearner` to estimate the joint distribution of confounders 
nonparametrically? 

The conditional survival functions are estimated by first estimating the 
conditional hazard, and then transforming into the conditional survival 
function. For a thorough explanation of these procedure see Section 3 from 
[@cai2019one]. Currently, `MOSS` requires the user to do this one-to-one 
tranformation from the conditional hazard to the conditional survival 
probabilities by calling the `hazard_to_survival()` method. We will address 
this later on.

Back to `initial_sl_fit()` -- we have the option to specify the following 
arguments in addition to the required data arguments. 

* `sl_failure`: SuperLearner library for failure event hazard, default = 
  c("SL.glm")
* `sl_censoring`: SuperLearner library for censoring event hazard, default = 
  c("SL.glm")
* `sl_treatment`: SuperLearner library for propensity score, default = 
  c("SL.glm")
* `gtol`: treshold for truncating propensity scores, default = 0.001)

It is highly recommended that you specify a more complex library than the 
default. The SuperLearner library arguments take a vector of strings 
corresponding to learners available in the `SuperLearner` `R` package: 
https://github.com/ecpolley/SuperLearner/tree/master/R. 

We do not review the `SuperLearner` `R` package in these workshops, but a handy 
tutorial crafted a few years ago by our colleague Chris Kennedy is freely 
available: [Guide to 
SuperLearner](https://cran.r-project.org/web/packages/SuperLearner/vignettes/Guide-to-SuperLearner.html).

```{r moss-2, message=FALSE, warning=FALSE}
# recall treatment was randomized
SL.ranger.faster = function(...) {
  SL.ranger(..., num.trees = 100)
}
sl_lib_decent <- c("SL.mean", "SL.glm", "SL.step.forward", "SL.bayesglm", 
                   "SL.ranger.faster", "SL.gam")

initial_fit <- initial_sl_fit(T_tilde, Delta, A, W, t_max, 
                              sl_censoring = sl_lib_decent, 
                              sl_failure = sl_lib_decent)

names(initial_fit)
```

The `initial_fit` object contains the fitted conditional densities for the 
failure events (`density_failure_1` for test treatment group, 
`density_failure_0` for standard treatment group), censoring events 
(`density_censor_1` and `density_censor_0` for test treatment and standard 
treatment groups, respectively), and propensity scores (a vector `g1W`).

The `density_failure_1` and `density_failure_0` both need to go through the 
`hazard_to_survival` method which populates the `survival` attribute of the 
object.

```{r, moss-2-transform, warning=FALSE, message=FALSE}
initial_fit$density_failure_1$hazard_to_survival()
initial_fit$density_failure_0$hazard_to_survival()
```

### 3. Perform TMLE adjustment of the initial conditional survival estimate {-}

The one-step TMLE is carried out by many local least favorable submodels (LLFMs) 
(performed via logistic regressions) with small step sizes. The one-step TMLE 
updates in small steps locally along LLFM, ensuring that the direction of the 
update is optimal around the current probability density. This procedure 
permits updates to the conditional hazard for all points on the survival curve 
(or any high-dimensional parameter in general), so that the conditional hazard 
can be transformed into a monotone survival curve after the algorithm.

At this point we are nearly ready to update the hazard using this constrained 
step size update. The only additional argument we need to define is

* `k_grid`: the vector of interested time points for estimation of the 
  conditional survival probability.

```{r moss-3-kgrid, message=FALSE, warning=FALSE}
k_grid <- 1:max(T_tilde)
initial_fit$density_failure_1$t <- k_grid
initial_fit$density_failure_0$t <- k_grid
```

We perform the TMLE adjustment of the initial conditional survival estimate by 
first creating a `MOSS_hazard` or `MOSS_hazard_ate` object and then calling 
the `iterate_onestep()` method for this object. `MOSS_hazard` and 
`MOSS_hazard_ate` correspond to different estimators:

* `MOSS_hazard`: one-step TMLE of the treatment-specific survival curve, and 
* `MOSS_hazard_ate`: one-step TMLE of ATE on survival probabilities

**One-step TMLE of the Treatment-Specific Survival**

```{r moss-3-standard, warning=FALSE, message=FALSE}
# estimate survival curve for standard treatment group
hazard_fit_standardA <- MOSS_hazard$new(
  A,
  T_tilde,
  Delta,
  density_failure = initial_fit$density_failure_0,
  density_censor = initial_fit$density_censor_0,
  g1W = initial_fit$g1W,
  A_intervene = 1,
  k_grid
)
psi_standardA <- hazard_fit_standardA$iterate_onestep()
```
```{r moss-3-test, warning=FALSE, message=FALSE}
# estimate survival curve for test treatment group
hazard_fit_testA <- MOSS_hazard$new(
  A,
  T_tilde,
  Delta,
  density_failure = initial_fit$density_failure_1,
  density_censor = initial_fit$density_censor_1,
  g1W = initial_fit$g1W,
  A_intervene = 2,
  k_grid
)
psi_testA <- hazard_fit_testA$iterate_onestep()
```


**One-step TMLE of ATE on Survival**
```{r moss-3-ate, warning=FALSE, message=FALSE}
hazard_fit_ate <- MOSS_hazard_ate$new(
  A,
  T_tilde,
  Delta,
  density_failure = initial_fit$density_failure_1,
  density_censor = initial_fit$density_censor_1,
  density_failure_0 = initial_fit$density_failure_0,
  density_censor_0 = initial_fit$density_censor_0,
  initial_fit$g1W
)
psi_ate <- hazard_fit_ate$iterate_onestep()
summary(psi_ate)
```

### 4. Compute standard error estimates for simultaneous inference {-}

The simultaneous confidence bands for the survival curve estimates can be 
constructed based on asymptotic linearity of the TMLE uniform in all time points 
considered. 

After creating a vector of survival curve estimates by defining a new 
`survival_curve` object, we will estimate the efficient influence curve 
using the TML-estimates generated in the previous step. We obtain estimates for 
all times specified in `k_grid` by calling the `all_t()` method for `eic` 
objects. The eic estimates are obtained for each subject and time, so we 
construct a matrix of these estimates and calculate a simultaneous confidence 
band with this matrix by calling the method `compute_simultaneous_ci()`. See the 
manuscript accompanying the `MOSS` package for more details on constructing 
simultaneous inference [@cai2019one].  

```{r moss-4-standard}
# estimate and obtain inference for survival curve for standard treatment group
survival_standardA <- survival_curve$new(t = k_grid, survival = psi_standardA)
survival_curve_standardA <- as.vector(survival_standardA$survival)

eic_standardA <- eic$new(
  A = A,
  T_tilde = T_tilde,
  Delta = Delta,
  density_failure = hazard_fit_standardA$density_failure,
  density_censor = hazard_fit_standardA$density_censor,
  g1W = hazard_fit_standardA$g1W,
  psi = survival_curve_standardA,
  A_intervene = hazard_fit_standardA$A_intervene
)
eic_matrix_standardA <- eic_standardA$all_t(k_grid = k_grid)
std_err_standardA <- compute_simultaneous_ci(eic_matrix_standardA)
upper_bound_standardA <- survival_curve_standardA + (1.96*std_err_standardA)
lower_bound_standardA <- survival_curve_standardA - (1.96*std_err_standardA)

plotdf_standardA <- data.frame(time = k_grid, est = survival_curve_standardA, 
                               upper = upper_bound_standardA, 
                               lower = lower_bound_standardA, 
                               type = rep("standard", length(k_grid)))

plot_standardA <- ggplot(data = plotdf_standardA, aes(x = time, y = est)) + 
  geom_line() +
  geom_ribbon(data = plotdf_standardA, aes(ymin = lower, ymax = upper), 
              alpha = 0.5) +
  ggtitle("Treatment-Specific Survival Curves Among Standard Treatment Group in
           Veterans’ Administration Lung Cancer Trial")
plot_standardA
```

```{r moss-4-test}
# estimate and obtain inference for survival curve for test treatment group
survival_testA <- survival_curve$new(t = k_grid, survival = psi_testA)
survival_curve_testA <- as.vector(survival_testA$survival)

eic_testA <- eic$new(
  A = A,
  T_tilde = T_tilde,
  Delta = Delta,
  density_failure = hazard_fit_testA$density_failure,
  density_censor = hazard_fit_testA$density_censor,
  g1W = hazard_fit_testA$g1W,
  psi = survival_curve_testA,
  A_intervene = hazard_fit_testA$A_intervene
)
eic_matrix_testA <- eic_testA$all_t(k_grid = k_grid)
std_err_testA <- compute_simultaneous_ci(eic_matrix_testA)
upper_bound_testA <- survival_curve_testA + (1.96*std_err_testA)
lower_bound_testA <- survival_curve_testA - (1.96*std_err_testA)

plotdf_testA <- data.frame(time = k_grid, est = survival_curve_testA, 
                           upper = upper_bound_testA, lower = lower_bound_testA, 
                           type = rep("test", length(k_grid)))

plot_testA <- ggplot(data = plotdf_testA, aes(x = time, y = est)) + 
  geom_line() +
  geom_ribbon(data = plotdf_testA, aes(ymin = lower, ymax = upper), alpha = 0.5) +
  ggtitle("Treatment-Specific Survival Curves Among Test Treatment Group in
           Veterans’ Administration Lung Cancer Trial")
plot_testA
```


```{r moss-4-ate-broken, eval=FALSE, echo=FALSE}
# estimate survival curve for ate
survival_ate <- survival_curve$new(t = k_grid, survival = psi_ate)
survival_ate$display(type = 'survival')

eic_ate <- eic$new(
  A = A,
  T_tilde = T_tilde,
  Delta = Delta,
  density_failure = hazard_fit_ate$density_failure,
  density_censor = hazard_fit_ate$density_censor,
  g1W = hazard_fit_ate$g1W,
  psi = survival_curve_ate,
  A_intervene = NULL
)
# TO-DO: fix simultaneous inference for ATE
eic_matrix_ate <- eic_ate$all_t(k_grid = k_grid)
std_err_testA <- compute_simultaneous_ci(eic_matrix_testA)
upper_bound_testA <- survival_curve_testA + 1.96 * std_err_testA
lower_bound_testA <- survival_curve_testA - 1.96 * std_err_testA

plotdf_testA <- data.frame(est = survival_curve_testA, 
                               upper = upper_bound_testA, 
                               lower = lower_bound_testA, 
                               time = k_grid,
                               type = rep("test", length(k_grid)))

plotdf <- rbind.data.frame(plotdf_standardA, plotdf_testA)
```
